---
title: "CodelistGenerator Extracting Procedure Codes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CodelistGenerator_extracting_procedure_codes}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
rm(list = ls()) # clear environment

library(here)
library(readr)
library(DBI)
library(RSQLite)
library(here)
library(dplyr)
library(stringr)
library(DT)
library(kableExtra)
library(tictoc)
library(CodelistGenerator)
```

## Creating a codelists for procedure
For this vignette we are going to create codelists for the procedure "colonoscopy" (https://athena.ohdsi.org/search-terms/terms/4249893). In these example we are going to generate a candidate codelists only looking for codes in the procedure domain.

## Getting the OMOP CDM vocabularies
As stated in other vignettes if you do not have ready access to a database with data in the OMOP CDM format (which will necessarily contain the vocabulary tables), you will first need to obtain the OMOP CDM vocabularies from https://athena.ohdsi.org. Once these are downloaded, we can make a vocabulary only database like so:

```{r, eval=FALSE}
library(readr)
library(DBI)
library(RSQLite)
library(here)
```

```{r, eval=FALSE}
#vocab.folder<-Sys.getenv("omop_cdm_vocab_path") # path to directory of unzipped files

vocab.folder<-"C:/Users/dnewby/Desktop/Athena Vocabs"

concept<-read_delim(paste0(vocab.folder,"/CONCEPT.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)
concept_relationship<-read_delim(paste0(vocab.folder,"/CONCEPT_RELATIONSHIP.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE) 
concept_ancestor<-read_delim(paste0(vocab.folder,"/CONCEPT_ANCESTOR.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)
concept_synonym<-read_delim(paste0(vocab.folder,"/CONCEPT_SYNONYM.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)
vocabulary<-read_delim(paste0(vocab.folder,"/VOCABULARY.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)

db <- dbConnect(RSQLite::SQLite(), ":memory:")
dbWriteTable(db, "concept", concept, overwrite=TRUE)
dbWriteTable(db, "concept_relationship", concept_relationship, overwrite=TRUE)
dbWriteTable(db, "concept_ancestor", concept_ancestor, overwrite=TRUE)
dbWriteTable(db, "concept_synonym", concept_synonym, overwrite=TRUE)
dbWriteTable(db, "vocabulary", vocabulary)
rm(concept,concept_relationship, concept_ancestor, concept_synonym, vocabulary)
vocabulary_database_schema<-"main"
```

The structure of each of these tables is described in detail at: https://ohdsi.github.io/CommonDataModel/cdm53.html#Vocabulary_Tables 

## Codelist from "colonoscopy"and its descendants
The simplest approach to identifying potential codes is to take a high-level code and include all its descendants. 
```{r, eval=FALSE}
library(dplyr)
library(stringr)
library(DT)
library(kableExtra)
```

```{r, eval=FALSE}

tic()
codes_from_descendants<-tbl(db, 
  sql(paste0("SELECT * FROM ",
     vocabulary_database_schema,
     ".concept_ancestor"))) %>% 
  filter(ancestor_concept_id=="4249893") %>% 
  select("descendant_concept_id") %>% 
  rename("concept_id"="descendant_concept_id") %>% 
  left_join(tbl(db, sql(paste0("SELECT * FROM ",
     vocabulary_database_schema,
     ".concept"))))%>%
  select("concept_id", "concept_name", 
         "domain_id", "vocabulary_id") %>% 
  collect()

toc()

saveRDS(codes_from_descendants, here("vignettes","proc_data_01.RData"))

```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
codes_from_descendants<-readRDS(here("vignettes","proc_data_01.RData"))
```

```{r,  message=FALSE, warning=FALSE }  
datatable(codes_from_descendants,
          rownames=FALSE,
          options = list(pageLength = 10,
                         lengthMenu = c(10, 20,50)))

```

## Generating a candidate codelist using Codelist Generator
To try and include all such terms that could be included we can use CodelistGenerator.

```{r, eval=FALSE }
library(CodelistGenerator)
```

First, let's do a simple search for a single keyword of "colonoscopy" without any exclusions, and including descendants but not ancestors.
```{r, eval=FALSE }
tic()
colonoscopy_codes1<-get_candidate_codes(keywords="colonoscopy",
                    domains="Procedure",
                    search_synonyms = TRUE,
                    fuzzy_match = FALSE,
                    exclude = NULL,
                    include_descendants = TRUE,
                    include_ancestor = FALSE,
                    verbose = TRUE ,
                    db=db,
                    vocabulary_database_schema =  vocabulary_database_schema)
toc()


#123.53 sec elapsed

```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
colonoscopy_codes1<-readRDS(here("vignettes","proc_data_02.RData"))
```

```{r,  message=FALSE, warning=FALSE }
datatable(colonoscopy_codes1,
          rownames=FALSE,
          options = list(
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```

What is the difference between this codelist and the one from 4249893 and its descendants?
```{r,  eval=FALSE }
code_comparison<-full_join(codes_from_descendants  %>% mutate(type1="4249893 and descendants"), 
                     colonoscopy_codes1  %>% mutate(type2="CodelistGenerator")) %>% 
  mutate(type=ifelse(!is.na(type1) & !is.na(type2),
                     "Both",
              ifelse(!is.na(type1) & is.na(type2),
                     "Only 4249893 and descendants",
              ifelse(is.na(type1) & !is.na(type2),
                     "Only CodelistGenerator",
                     NA)))) %>% 
  select(-c("type1","type2"))
```

